Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-09-11 SupportingInformation.Rmd
modified@sept10

@@

```{r tab_soilCN_asym,echo=F,results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
knitr::kable(tab_soilCN_asym, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption = "**Soil CN data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis. Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (see \\nameref{MT-NPA} in the manuscript). One thousand Kendall-preserving and 1000 Spearman-preserving surrogates with normal copula structure were created, and the same statistics were computed for the surrogates. A table entry $<N$ indicates the value of the given statistic on the data was less than its value on $N$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
#kable_styling(, position = "float_left")
```

@@

@@

\newpage
```{r tab_stat_npa_asym_2,echo=F,results="markup"}
temp<-make_table_stat_fn(data=tab_asym[[2]])
knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[2]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_2}"))
```

\newpage
```{r tab_stat_npa_asym_3,echo=F,results="markup"}
temp<-make_table_stat_fn(data=tab_asym[[3]])
knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[3]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_3}"))
```

@@

## Need to think about mechanisms for biogeochemical, evolutionary, and community data

@@

>>>>

# Description of Data \label{Data_detailed}

For bivariate analysis we used four datasets which are briefly described as follows;
First, to explore the dependence of **soil organic C** and total **soil N** across a wide geographic range (conterminous US), ecosystem type, and traceable methodology we used the RaCA (Rapid Carbon Assessment)
database [@Wills2014, **or Wills and Loecke pending???**]. Specifically, the soil organic C and total soil N
stocks (Mg C or N m$^{-2}$ ) to 1m of soil depth from the central pedon of the 5930 sites were included in this
study.

Next, the [**BMR data**](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) contains the trait contrasts from phylogenetically independent contrasts [@SukumaranH2010] of data reported by [@GenoudIM2017] on the basal metabolic rate (BMR) and mass of mammals species. Specifically, the "SELECT" BMR data in their database were used for those species for which data were available. A phylogeny in NEXUS [@NEXUS] format was extracted from the supplementary information of [@GenoudIM2017]. Version 4.4.0 of the DendroPy library [@SukumaranH2010] was used to calculate phylogenetic independent contrasts using Python 2.7. For each contrast, the path length spanned by the contrast (sum of the branch lengths) and the contrast for the mass, natural logarithm of the BMR and natural logarithm of the mass were written as a row in the [bmr-contrasts.csv](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) file.

[**Third dataset**](https://github.com/sghosh89/BIVAN/blob/master/Data/Anolis/contrasts.csv) was a raw data on measurements of anatomical features of several species of *Anolis* lizards, which was downloaded
from the Dryad data package [@MahlerIRL2013Data] associated with [@MahlerIRL2013].
The [contrasts.csv file](https://github.com/sghosh89/BIVAN/blob/master/Data/Anolis/contrasts.csv) was produced by using version 4.4.0 of the DendroPy library [@SukumaranH2010]  to calculate phylogenetic independent contrasts on the the data in the GA_Anolis_traits.csv file of Mahler et at. using their maximum clade credibility tree (GA_Anolis_MCC.nex) as the phylogeny.

Absolute values of traits for each contrast are considered to construct corresponding copula for second and third datasets.

Fourth, we used **Ceder-creek data for e120** which resulted from a biodiversity experiment, carried out by Tilman and his co-warkers [@Tilman2001; @Tilman2006] over many years to collect [plant above ground biomass](https://www.cedarcreek.umn.edu/research/data/dataset?ple120) and [plant species percent cover](https://www.cedarcreek.umn.edu/research/data/dataset?pce120) data. These datasets were used to calculate plant-aboveground biomass and Shannon's diversity index (H), averaged over each plot (9m by 9m),
for year 2000.

<!--The raw data and copula for each four datasets are shown along each column of Figure \ref{MT-fig_biv_raw_cop}.-->

Similarly, we applied multivariate analysis on four datasets; first two data sets [for aphid-count and aphid-firstflight date] were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014]. These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species. We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony. In our case, we chose species=10 [Name : "Green spruce aphid"] for aphid-abundance data and species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data and considered only those locations which had atleast 30 years of data (see Figure \ref{fig_aphid_map}).

<!--Figure \ref{MT-fig_aphid_raw_cop} represents raw and copula data plot for aphid count ("Green spruce aphid") and aphid-firstflight date ("Leaf-curling plum aphid") taken from two locations.-->

Our third dataset for multivariate analysis is a plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species and for 56 years (1958-2013). These data were obtained from the Continuous Plankton Recorder (CPR) data set and maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS). The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
location in the ocean. Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979]. All the samples available for a given variable in a given area in a given month of a given year were
averaged, producing a monthly time series for each year. Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available). Monthly values were converted to yearly values by averaging over all months. This approach was used by [@Sheppard2017] earlier. But here we only allowed maximum 4 months in a year with missing data which were replaced by those medians. If a year got 5 or more missing monthly data value then we set NA value for that particular year. We used data for species=16 [Name : "Ceratium furca"] which is a dicoflagellate commonly found in UK seas. Among 26 locations, only those having atleast 45 years time series data, were considered (see Figure \ref{fig_plankton_map}).

Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the
Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio. Additional information on the history of the site and soil types is detailed in @Jarecke2016. Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at
daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
additional details???** on greenhouse gas flux methods for this site. We threw out all locations that had fewer than 50 dates data.

# Nonparametric analysis of tail dependence \label{NPA}

Left (or lower)- and right (or upper)-tail dependence quantify the extent to which two ranked
variables $u_i$, $v_i$ for $i=1,...,n$ 
(see \nameref{MT-Model_selection} for defintions of $u_i$ and $v_i$ in the manuscript)
are related to each other in the left and right tails, respectively, of their 
distributions [@nelsen2007_copula, @joe2014_dependence]. 
Our nonparametric study of tail dependence uses three
statistics which can quantify the extent to which $u_i$ and $v_i$ are 
related in any part of their distributions. We here describe the 
statistics. The statistics were formulated with positively associated
variables in mind. All the variables we study are positively associated
when they are significantly associated. So definitions and associated 
diagrams will assume positive associations, though generalizations are 
possible. Given two bounds 
$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
fall (these are the blue and red lines, respectively, in 
Fig. \ref{fig_stat_image}). Our statistics quantify the dependence
between $u_i$ and $v_i$ in the region bounded by these lines.
Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
dependence in the left parts of the distributions of $u$ and $v$, and 
using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
the right parts of the distributions.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
\caption{Schematic diagram supporting defintions of nonparametric tail dependence statistics \label{fig_stat_image}}
\end{center}
\end{figure}

The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
correlation of $u_i$ and $v_i$ that is attributable to the points 
lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$, i.e.,
\begin{equation}\label{eq.Cor}
   \cor_{l_b,u_b}(u,v) = \frac{1}{(n-1)\sqrt{\var(u)\var(v)}}\sum 
   (u_i-\mean(u)) (v_i-\mean(v)),
\end{equation}
where means and variances are computed using all $n$ data 
points, but the
sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
$u_i+v_i < 2u_b$. 
Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
between $u$ and $v$ in the region given by the bounds.

We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
$u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
and a distance less than $h$ from the line $v=u$, divided by the 
total number of points within the bounds. This function is
defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
is half the longer of the two segments obtained by intersecting the
bounds with the unit square. It is easy to see $S(0)=0$ and 
$S(h_{\text{max}})=1$. We defined $S_i(h)$, for the same range 
of $h$, to be the area within the bounds and the unit square
and within distance $h$ of the line $v=u$, divided by the area
within the bounds and the unit square. This is the expected value of
$S(h)$ for independent data. We then defined
\begin{equation}\label{eq.P}
   P_{l_b,u_b} = \int_0^{h_{max}} ( S(h) - S_i(h) ) dh.
\end{equation}
As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
indicate stronger positive association
between $u$ and $v$ in the region given by the bounds.

The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
distance between points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ and the line $v=u$. Unlike $\cor_{l_b,u_b}$ and 
$\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dsq_{l_b,u_b}$ is 
then small. 
Thus large values of $\cor_{l_b,u_b}$ and $\Ps_{l_b,u_b}$ indicate strong 
dependence in the portions of the distributions of $u$ and $v$ given by 
the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
$\Dsq_{l_b,u_b}$ indicate strong dependence.

For large datasets (i.e., datasets with large $n$),
we can consider $l_b$ and $u_b$ close together without incurring undue
sampling variation in our statistics, and we can consider 
several different bands $(l_b,u_b)$ to understand how dependence
varies in different parts of the distributions. But for datasets with 
smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
We use the shorthand $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, and $\Dsq_u=\Dsq_{0.5,1}$.

The difference $\cor_l-\cor_u$ is expected to be positive if 
dependence in the left halves of the distributions is stronger 
than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
This holds, for instance, for the Clayton copula [@nelsen2007_copula].
The difference $\Dsq_u-\Dsq_l$, in which the "upper" and
"lower" statistics occur in opposite order, is also expected to be 
positive under the same conditions. We use these three differences
to detect asymmetry of tail dependence. Likewise, for large datasets,
we consider $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.

We tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulas (for details see Appendix \ref{Test_npa_stat}).

For bivariate datasets, we compared values 
of the statistics $\cor_{l_b,u_b}$,
$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to 
distributions of values of the same statistics computed on surrogate datasets that were produced from 
the empirical data by randomizing it in a special way. The surrogate/randomized datasets had exactly 
the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
correlation (the algorithm had two versions, one for preserving each correlation coefficient), 
but had Normal copula structure. Thus our comparisons tested the null hypothesis 
that our statistics took values on the empirical data no different from what would have been expected 
if the copula structure of the data were Normal, but the data were otherwise 
statistically unchanged.
Details of how surrogates were prroduced are in Appendix \ref{surrog_test}, and code is in [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R).

The comparison of one of the
statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
empirical data, to the distribution
of its values computed on many surrogate datasets 
provides a test of whether dependence in the part of the distribution specified 
by $l_b$ and $u_b$ is different from
what would be expected from a null hypothesis of normal copula structure but 
unchanged marginals and unchanged 
Kendall (or Spearman)
correlation. Thus the comparison helps answer the first two parts of question \hyperref[MT-Intro]{\textbf{(Q1)}} 
from the \nameref{MT-Intro} for bivariate datasets. 
Significant deviations correspond to deviations from normal 
copula structure. In particular, deviations 
using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
different from that of a normal copula;
likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
from that of a normal copula.

The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, 
as computed on the 
empirical data, to the distribution
of its values on many surrogate datasets 
provides a test of asymmetry of tail dependence. Thus this comparison helps answer the
third part of question \hyperref[MT-Intro]{\textbf{(Q1)}} from the \nameref{MT-Intro} for bivariate datasets.

For multivariate datasets, a spatial resampling scheme was used instead of the surrogate 
approach described above.
The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models. For each pair of 
locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
values $s_{ij}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the 
average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a 
distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
revealed by these confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
Normal null hypothesis, and answering question \hyperref[MT-Intro]{\textbf{(Q1)}} from the \nameref{MT-Intro}.
Bootstrapping was performed by resampling the locations of measurement, with replacement. 
This corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
[@bjornstadfalck2001 ; @walteretal2017].

@@

```{r tab_soilCN_asym,echo=F,results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
knitr::kable(tab_soilCN_asym, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption = "**Soil CN data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis. Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (see \\ref{NPA}). One thousand Kendall-preserving and 1000 Spearman-preserving surrogates with normal copula structure were created, and the same statistics were computed for the surrogates. A table entry $<N$ indicates the value of the given statistic on the data was less than its value on $N$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
#kable_styling(, position = "float_left")
```

@@

# Cause of non-Normal copula structure : the Moran effect and copula structure in populations \label{Cause_Moran}

To begin with, we chose a two habitat patch ($i = 1, 2$) with AR(1) model [Eq. \ref{eq.AR1}] where population $P_{i}$ at time $(t+1)$ depends upon its past value at $(t)$ and the iid noise (assumed to be caused by environmental fluctuations) $\epsilon_{i} (t)$.

\begin{equation}\label{eq.AR1}
P_{i}(t+1)=\beta P_{i}(t)+\alpha \epsilon_{i}(t)
\end{equation}

To keep $Var(P_{i}(t+1))$ equal to unity, we set noise coefficient $\alpha$ in terms of AR coefficient $\beta$ as $\alpha = \sqrt(1-\beta^2)$. We were interested to explore the basic idea of Moran's theorem that if noise $\epsilon_{1}(t)$ and $\epsilon_{2}(t)$ at two different habitat patches had some specific assymetric tail dependence and then what will be the outcome dependence in the populations at those habitat patches. Should their dependence be exactly the same as of the noises or not ?

To test that our approach is of two way : first, we calculated the parameters of the noise-copula (for example, we chose Clayton with lower tail dependence and Survival Clayton with upper tail dependence here) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively. Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome which was a bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig. \ref{fig_Cause4copula_params_GOF_500}). We also tested the goodness of fit of the estimated parameters for population-copula and its standard error due to sample size and reported p-values for both statistics (Cramer-von Mises and Kolmogorov-Smirnov). p-values greater or equal to $0.05$ (above the dashed line in Fig. \ref{fig_Cause4copula_params_GOF_500}) were considered as a good fit of estimated parameters. Fig. \ref{fig_Cause4copula_params_GOF_2500} represents similar feature as of Fig. \ref{fig_Cause4copula_params_GOF_500} but here we considered last $2500$ simulated population data points for our parameters-estimation, thus making our comparison more sensitive on basis of p-value threshold at $0.05$. (For code see [`Plotter_Cause4copula_GOF`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).

Secondly, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric statistics (see \ref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations. Then we plotted mean of each measured quantity with standard error against Spearman's Rho (or Kendall's Tau) of the noise copula family. Fig. \ref{fig_Cause4copula_C_spear} and Fig. \ref{fig_Cause4copula_C_kend} represent this picture for noise generating from a lower tail dependent Clayton copula whereas for Fig. \ref{fig_Cause4copula_SC_spear} and Fig. \ref{fig_Cause4copula_SC_kend} noise came from a upper tail dependent Survival Clayton family. We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach, here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which ensured the Moran effect on the population data caused by environmental noise. (For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).

@@

```{r tab_stat_npa_asym_2,echo=F,results="markup"}
temp<-make_table_stat_fn(data=tab_asym[[2]])
knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[2]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_2}"))
```

```{r tab_stat_npa_asym_3,echo=F,results="markup"}
temp<-make_table_stat_fn(data=tab_asym[[3]])
knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[3]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_3}"))
```

@@

# Cause of non-Normal copula structure : Asymmetric sensitivity and copula structure in populations \label{Cause_asym_sens}

To test the assymetric sensitivity, we constructed a two habitat patch model for population $P_{i} (i=1,2)$ as follows;

\begin{equation}\label{eq.AS}
P_{i}(t+1) = b P_{i}(t) + g(\epsilon_{i}(t)) + h(\delta_{i}(t))
\end{equation}

$b$ was the linear growth rate. Global environmental fluctuation function $g(\epsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq. \ref{eq.AS_g}) where $\epsilon=(\epsilon_{1}, \epsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\epsilon}$.

\begin{align}
g(\epsilon) = \begin{cases} 
g_{1} = \begin{cases}
\epsilon & \text{if $\epsilon< 0$}; \\
0 & \text{otherwise}.
\end{cases}\\ \\
g_{2} = \begin{cases}
\epsilon & \text{if $\epsilon> 0$}; \\
0 & \text{otherwise}\label{eq.AS_g}.
\end{cases}
\end{cases} \text{where } \epsilon = N(\textbf{0},\Sigma_{\epsilon})\;\ \text{with} \;\ \Sigma_{\epsilon} = \begin{bmatrix} 1 & r \\ r & 1 \end{bmatrix}
\end{align}

Local environmental fluctuation function $h(\delta)$ varied according to Eq. \ref{eq.AS_h} where $a$ was the scaling parameter between $0$ to $1$. $\delta=(\delta_{1}, \delta_{2})$ was an iid across $i=(1,2)$ and time with mean zero and covariance matrix $\Sigma_{\delta}$.

\begin{equation}\label{eq.AS_h}
h(\delta) = a \delta = aN(\textbf{0},\Sigma_{\delta})\;\ \text{with} \;\ \Sigma_{\delta} = \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}
\end{equation}

We simulated our model (\ref{eq.AS}) for different values of $b$ with $a=0.2$, $r=0.8$. $10000$ points were generated from each local and global fluctuations to apply on the initial populations and then after simulation we kept only last $1000$ time step of simulated points to plot the population-copula for a specific $b$ and a specific functional form of $g$  (For code see [`Simulator_asym_sens`](https://github.com/sghosh89/BIVAN/blob/master/asym_sensitivity.R) function). Fig. \ref{fig_ms_asym_b_0.1} and Fig. \ref{fig_ms_asym_b_0.9} show these results for $b=\pm0.1$ and $b=\pm0.9$, respectively. We also carried out model selection and non-parametric statistical analysis with each copula of those figures and summarized our results in Tables \ref{tab_summary_stat_asym_sens} - \ref{tab_stat_npa_asym_8}. Table \ref{tab_summary_stat_asym_sens} represent the model selection results with AIC weights for best fitted copula and normal copula, model averaged lower and upper tail dependence, p-values (both Cramer-von Mises statistic and Kolmogorov-Smirnov statistic) for goodness of fit test. In each case, successful bootstrap was $100\%$. According to our model selection approach, for $b=\pm0.1$, non-normal copula (with higher AIC weights) were always best-fit with some asymetric tail dependence whereas for $b\pm0.9$, most of the time normal was best. It is probably because the populations get long-time exposure to environmental changes for slower growth(or decay) rate and so global fluctuation can affect them more efficiently. Our non-parametric analysis results (Tables \ref{tab_stat_npa_asym_1} -\ref{tab_stat_npa_asym_8}) are consistent with this model selection approach.

#  Need to think about mechanisms for biogeochemical, evolutionary, and community data

@@

\begin{figure}[!h]
\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.1 cm} (D)} \\
\vspace{-0.5 cm} 
\begin{center}
\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC.pdf}\\
\end{center}
\textbf{ \hspace{3.1 cm} (E) \hspace{2.4 cm} (F) \hspace{2.3 cm} (G) \hspace{2.3 cm} (H) \hspace{2.3 cm} (I)}\\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_5stats.pdf}
\caption[Spatial Taylor's Law results]{Spatial Taylor's Law results: log(spatial variance) vs. log(spatial mean) for $ns=10$ number of (A) Clayton, (B) Normal, (C) Frank, (D) Survival Clayton surrogates; (E) p-value for test of linearity, (F) p-value for test of hhomoskedasticity, (G) root mean square error, (H) intercept, (I) slope of the first panel plots. \label{fig_spTL}}
\end{center}
\end{figure}

# Consequence of non-Normal copula structure : effect on skewness of the spatial mean \label{Consequence_skewness}

Because of overall distinct tail dependence of the multivariate copula structure, the skewness of the spatially-averaged time-series data, changed significantly from the individual skewness of time-series data. Spatial mean of the time series data extracted that tail dependence from the multivariate data set, if any, and reflected in the averaged skewness. We tested null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula. So, first we generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same. Then we calculated the skewness of spatially averaged surrogate time series data, as we did with the actual multivariate data set (For code see [skewness_multivar_dataset.R](https://github.com/sghosh89/BIVAN/blob/master/skewness_multivar_dataset.R). We made a histogram of surragate skewness and found if actual skewness were within $95\%$ quantile or not. If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness).

To verify that we used our four multivariate datasets : aphid-count data, aphid-firstflight data, plankton data and methane-flux data. Figure \ref{fig_skewness_count} showed that averaged skewness was much lower than expected if the aphid count data came from a normal copula. In that example, $p$ value was denoted by the fraction of surrogate-skewness which were lower than the averaged skewness of the given data (red line in Figure \ref{fig_skewness_count}). As $p<0.05$, we rejected $H_{0}$ in support of significantly stronger lower tail dependence which was also consistent with the non-parametric analysis and model fitting results. On the other hand, Figure \ref{fig_skewness_ff} demonstrated that averaged skewness was much higher than expected if the aphid first flight data came from a normal copula. As $p$ value (the fraction of surrogate-skewness which were greater than the averaged skewness of the given data - red line in Figure \ref{fig_skewness_ff}) was less than $0.05$, we reject $H_{0}$ in support of upper tail dependence. Figure \ref{fig_skewness_pns} showed that averaged skewness was lower than expected if the plankton data came from a normal copula. Our previous results on non-parametric analysis and model fitting also supported lower tail dependence. But, as $p>0.05$, we could not reject $H_{0}$ in that case. Finally, for methane-flux data, the averaged skewness was lower than expected if the data came from a normal copula (Figure \ref{fig_skewness_methane}). As $p<0.05$, we rejected $H_{0}$ and that result was consistent with the non-parametric analysis and model selection results which also favored lower tail dependence.

# Consequence of non-Normal copula structure : Effect on Taylor's law \label{Consequence_TL}

Spatial Taylor's law, which shows linear relationship between `log`(variance) and `log`(mean) of population densities of a species from different locations, is a well-known pattern found in ecology and related fields [@Taylor1961, @Taylor1988, @Cohen2013, @Doring2015, @Kuo2016, @Xu2016, @Reuman2017]. Here, we examined the nature of the slope and how it changed if population time series (taken as $50$ years) from different locations (say, $numloc=25$) had some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula. First, we simulated (N $\times$ n $\times$ nr) dimensional normal copula with covariance $\rho =0.7$, where $(N=50,n=25,nr=10)$ are the length of timeseries, number of sampling locations and number of replica for that copula, respectively. Then we generated Kendall-correlation preserving \textbf{C}layton, \textbf{F}rank and $\textbf{S}$urvival \textbf{C}layton surrogates with same dimension of \textbf{N}ormal. With these four different types of copula we computed `log`(spatial variance) and `log`(mean) and plotted in Fig.\ref{fig_spTL}(A-D). We also tested statistically that whether linearity, homoskedasticity is maintained in each copula dataset or not and plotted our results alongwith root mean square error, intercept and slope in Fig.\ref{fig_spTL}(E-I) (For code see [`copsurrognd.R`](https://github.com/sghosh89/BIVAN/blob/master/copsurrognd.R) and [`TaylorsLawTools.R`](https://github.com/sghosh89/BIVAN/blob/master/TaylorsLawTools.R)).

# Consequence of non-Normal copula structure : effect on extinction risk \label{Consequence_ext_risk}

In order to study the effects of non-Normal copula structure on extinction risk for metapopulations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;
<!--did I write correct spelling for lewton model?-->

\begin{equation}\label{eq.extrisk_model}
P(t+1) = \exp\left [  r\left ( 1-\frac{P(t)}{K} \right ) + \epsilon(t) \right ] P(t)
\end{equation}

$r$ is the liner growth rate and $K$ is the carrying capacity of the environment. $\epsilon$ is the environmental noise which might take stronger right tail (means stronger upper tail)  or stronger left tail (means stronger lower tail) dependence. We then considered each population $P(t+1)$ was updated depending upon $D$, the dispersal matrix, which could take either forms; one case when each population patch on a linear chain interacted with its two nearest neighbors (we called it local dispersion) or when any patch was allowed to interact with the rest of all (we called it as global dispersion). We simulated Eq. \ref{eq.extrisk_model} for $10000$ simulations and calculated extinction risk after $25$ years with different possibilities (for code see [mydispersal.R](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R)) and summarized our findings as follows;

- LC model (r=0) :

    * For LC model, as expected, when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one. [see first subplot of Fig. \ref{fig_risk_local_disp} and Fig. \ref{fig_risk_global_disp}]

- Ricker model (r $\neq$ 0):

    * For ricker model when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$. For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations. [see Fig. \ref{fig_risk_local_disp} - \ref{fig_risk_global_disp}]

    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_K_effect}]
